### 2023-04-11 txt/tmp.tex

With the surge of availability of quantum computing architectures in recent
years it becomes increasingly difficult to keep track of their relative
performance. To make this case even more difficult, various providers give
access to different figures of merit for their architectures. Our package allows
the user to test various architectures, available through \texttt{qiskit} and
Amazon BraKet using problems with simple operational interpretation. We provide
one example built-in in the package. Furthermore, we provide a powerful tool for
the users to extend the range of available problems in a way that suits their
needs.

Due to this possibility of extension, the users are able to test specific
aspects of their architecture of interest. For example, if their problem is
related to the amount of coherence (the sum of absolute value of off-diagonal
elements) of the states present during computation, they are able to quickly
prepare a custom experiment, launch it on desired architectures, gather the
result, based on which they can decide which specific architecture they should
use.

Finally, we provide the source code of PyQBench on GitHub~\cite{pyqbenchgithub} under an open
source license which will allow users to utilize and extend our package in their
specific applications.