### 2023-08-26 reviews/softwarex/reviews_1/response/response.tex
Add response to reviewer

\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{enumitem}

\newcommand{\1}{{\rm 1\hspace{-0.9mm}l}}

\newenvironment{response}{\vspace{1em}\noindent\textbf{Response}}{\vspace{1em}}

\begin{document}
We would like to thank the reviewers for their comments. Below we provide
a response to their specific remarks.

\section{Reviewer 1}

\begin{enumerate}
  \item The cross-entropy benchmarking technique used in the validation of the Sycamore-53 QPU in
    quantum supremacy experiments (https://www.nature.com/articles/s41586-019-1666-5) should also
    be mentioned in the Introduction (https://www.nature.com/articles/s41567-018-0124-x).

    \begin{response} We modified manuscript to include mention of cross-entropy benchmarking as
      suggested by the referee.
    \end{response}
  \item The generalization of the suggested benchmarking techniqiue to multiple qubits should be
    described in more detail, explicitly mentioning whether the complexity of the derivation and
    actual benchmarking depends on the number of qubits.

    \begin{response} We thank the referee for this suggestion. We extended the manuscript with a
      paragraph describing differences between a single--qubit and higher-dimensional cases.
    \end{response}

\end{enumerate}
\section{Reviewer 2}

\end{document}